---
title: "rain.hmm: Hidden Markov Models with Variational Bayes"
output: github_document
---

This package is designed...

# Variational Bayes Basic Function

The basic function uses Variational Bayes assuming exponential mixture components:


```{r example}
library(rain.hmm)
#Pulling the simplified Chesapeake Bay dataset included within the package as a tester.

set.seed(1)
data <- cbayrain[,-1]
data <- data[,1:10]

raindata <- array(NA, dim = c(1840, 1, 10))

for(i in 1:10){
  raindata[,,i] <- data[,i]
}

#the [-1] removes the dates column. Shrinking to 10 locations for ease of showcase.

#The VBEM function requires you to start with some potential priors, as well as pre-pick the number of days, years, mixtures, and number of locations you collected data from.

##Assuming 3 states, 3 mixtures and 10 locations, with 1840 days of data collected and not split into years

S = 3
M = 3
L = 10

##Initial estimates of the hyperparameters:
gamma_0     <- matrix(data = c(.5,2,1.5,9,2,16),nrow = S,ncol = M-1,byrow = T) # shape of exponential rate
gamma_0     <- array(gamma_0,dim = c(S,M-1,L))
delta_0     <- matrix(data = c(2,2,2,2,2,2),nrow = S,ncol = M-1,byrow = T) # rate of exponential rate
delta_0     <- array(delta_0,dim = c(S,M-1,L))
zeta_0      <- matrix(data = c(6,8,6,6,7,7,8,6,6),nrow = S,ncol = M,byrow = T)/(M-1) # Dirichlet prior parameters for mixing probabilities
zeta_0      <- array(zeta_0,dim = c(S,M,L))
alpha_0     <- matrix(rep(10,S^2),nrow = S,byrow = T)/S # Dirichlet prior for transition matrix rows
xi_0        <- rep(1,S)/S # Dirichlet prior for initial probabilities

param <- list('gamma_shape' = gamma_0, 'gamma_rate' = delta_0, 'zeta' = zeta_0, 'alpha' = alpha_0, 'xi' = xi_0)

Model <- fit.VBEM(D = 1840, S = 3, Y = 1, L = 3, M = 3, hypers = param, dist = "exp", data = raindata)
```

The output is a list, containing information such as the posterior hyperparameters, or the ELBO/DIC.

```{r posterior pi}
print(Model$posteriors$pi)
```
```{r posterior transition}
Model$posteriors$transmat
```
## Post-Formulation Functions

This package includes additional functions like the viterbi-encoding algorithm, so it does not have to be programmed on the side:

```{r viterbi}


```


# Stochastic Version
As data gets more complex, with more locations and possibly adding in more mixtures and states, it becomes necessary to create a less intensive method. This is the purpose of the stochastic version of the function.

```{r codycody, include = F}
library(rain.hmm)
library(Rmpfr)
```


```{r stochastic}
#Define some hyperparameter guesses
gamma_0     <- matrix(data = c(.5,2,1.5,5,2,10),nrow = 3,ncol = 2,byrow = T) # shape of exponential rate
gamma_0     <- array(gamma_0,dim = c(3,2,10))
delta_0     <- matrix(data = c(2,2,2,2,2,2),nrow = 3,ncol = 2,byrow = T) # rate of exponential rate
delta_0     <- array(delta_0,dim = c(3,2,10))
zeta_0      <- matrix(data = c(3,4,3,3,3.5,3.5,4,3,3),nrow = 3,ncol = 3,byrow = T) # Dirichlet prior parameters for mixing probabilities
zeta_0      <- array(zeta_0,dim = c(3,3,10))
alpha_0     <- matrix(rep(10,9),nrow = 3,byrow = T)/3 # Dirichlet prior for transition matrix rows
xi_0        <- c(.4,.3,.3) # Dirichlet prior for initial probabilities
param <- list('gamma_shape' = gamma_0, 'gamma_rate' = delta_0, 'zeta' = zeta_0, 'alpha' = alpha_0, 'xi' = xi_0)

#StoModel <- fit.VBEM(D = 92, S = 3, Y = 20, L = 10, M = 3, hypers = param, dist = "stochexp", data = raindata)
```

```{r result}
#print(StoModel$posteriors$pi)
#print(StoModel$iternum)
```


# Gamma Mixtures
